### 2023-02-27 chapter0/chapter0.tex
Removed tennis court analogy and replaced with statistic on volume

Some examples will illustrate the central role of proteins more clearly. Blood is a part of the circulatory system which is responsible for transporting nutrients and waste. Though blood is a complex mixture, containing a cocktail of cells, proteins, sugars, gases, and ions dissolved in a medium of water, its primary cellular component is red blood cells. These cells, which give blood its red color, ferry oxygen from lungs throughout the body. While water can dissolve some oxygen, the body requires more oxygen more quickly than is available in the aqueous component of blood alone. Thus, red blood are packed with a special protein called hemoglobin, which binds\footnote{In biology, ``binding'' is a slippery word whose exact meaning can varying greatly depending on the context. However, it generally means that a molecule physically interacts with another molecular for an extended period.} oxygen. Each red blood cell contains as many as 270 million molecules of hemoglobin, each of which can carry up to four oxygen molecules~\cite{Pierig2008}. To put this value in perspective, if each molecule of hemoglobin were the size of a tennis ball, the hemoglobin molecules in a single red blood cell would cover a tennis court in over 10 packed layers\footnote{A standard tennis ball is 1.35 in or 0.1125 ft in radius, so the volume of 270 million tennis balls of this size packed as efficiently as possible is given by $2.70 \times 10^6 \times \frac{4}{3} \pi (0.1125 \text{ ft})^3 \times \frac{3\sqrt{2}}{\pi} = 2.17 \times 10^4 \text{ ft}^3$. A singles tennis court is 78 ft by 27 ft or 2106 ft\textsuperscript{2}, so dividing the first by the second yields slightly over 10 layers.}. Because red blood cells are so dense with hemoglobin, any defect in hemoglobin can dramatically impact the structure of the red blood cells themselves. A well-studied example is sickle cell disease where an error in the body's hemoglobin molecules deforms red blood cells into a characteristic sickle shape. This prevents them from easily flowing through blood vessels, resulting in pain and oxygen deprivation.

>>>>

Some examples will illustrate the central role of proteins more clearly. Blood is a part of the circulatory system which is responsible for transporting nutrients and waste. Though blood is a complex mixture, containing a cocktail of cells, proteins, sugars, gases, and ions dissolved in a medium of water, its primary cellular component is red blood cells. These cells, which give blood its red color, ferry oxygen from lungs throughout the body. While water can dissolve some oxygen, the body requires more oxygen more quickly than is available in the aqueous component of blood alone. Thus, red blood are packed with a special protein called hemoglobin, which binds\footnote{In biology, ``binding'' is a slippery word whose exact meaning can varying greatly depending on the context. However, it generally means that a molecule physically interacts with another molecular for an extended period.} oxygen. Each red blood cell contains as many as 270 million molecules of hemoglobin, each of which can carry up to four oxygen molecules~\cite{Pierig2008}. Because red blood cells are so dense with hemoglobin, composing roughly 35\% of their total volume, any defect in hemoglobin can dramatically impact the structure of the red blood cells themselves~\cite{Kanias2009}. A well-studied example is sickle cell disease where an error in the body's hemoglobin molecules deforms red blood cells into a characteristic sickle shape. This prevents them from easily flowing through blood vessels, resulting in pain and oxygen deprivation.
